Particular examples of planar integral point sets and their classification \footnote{ This work was carried out at Voronezh State University and supported by the Russian Science Foundation grant 19-11-00197. }

### 2025-11-16 AvdMomZvolA_examples/ips_examples.tex
[AMZA] Abstract (Rus and Eng)

\paragraph{Abstract.}
	A planar integral point set is a set of non-collinear points in plane
	such that for any pair of the  points the Euclidean distance
	between them is integral.
	We discuss the classification of planar integral point sets
	and provide examples of sets that are not covered by the existent classification.

>>>>

\paragraph{Abstract.}

A planar integral point set (PIPS) is a finite set of non-collinear points in the Euclidean plane
such that the Euclidean distance between any pair of points is an integer.
These sets are characterized by their cardinality (finite number of points),
diameter (maximum pairwise distance), and characteristic (the smallest positive integer \(q\)
such that all triangular areas are commensurable with \(\sqrt{q}\)).
The characteristic remains invariant under translations, dilations, reflections, and even the addition or removal of points.

Existing classifications include sets in semi-general position (no three points collinear)
and general position (no three collinear and no four concyclic).
Circular sets and facher sets (all but one points on a line) are prominent examples,
but finding sets in general position is challenging; for instance, the largest known has seven points,
and no eight-point example exists yet.

This work introduces new examples to advance classification, including rails sets (points on two parallel lines),
sets with multiple symmetries.
We also present sets with many shared points that cannot be merged.
These constructions highlight potential of sequential extensions and limitations in merging sets,
offering insights into the structure and properties of planar integral point sets.

\paragraph{Аннотация.}

Плоское целоудалённое множество есть конечное множество точек на евклидовой плоскости,
не содержащееся ни в какой прямой,
такое что евклидово расстояние между любой парой точек является целым числом.
Эти множества характеризуются своей мощностью (конечным числом точек),
диаметром (максимальным попарным расстоянием) и характеристикой (наименьшим положительным целым числом \(q\),
таким что площади всех треугольников, образованных точками множества, соизмеримы с \(\sqrt{q}\)).
Характеристика инвариантна относительно при сдвига, растяжения, отражения, а также добавления или удаления точек.

Существующие классификации включают множества в полуобщем положении (никакие три точки не лежат на одной прямой)
и в общем положении (никакие три точки не лежат на одной прямой и никакие четыре не лежат на одной окружности).
Классическими примерами являются круговые множества и веерные множества (все точки, кроме одной, лежат на одной прямой),
однако нахождение множеств общего положения представляет значительные трудности;
например, наибольшее известное множество имеет семь точек, и пока не найдено множество из восьми точек общего положения.

В данной работе представлены новые примеры для развития классификации,
включая рельсовые множества (точки на двух параллельных прямых),
множества с несколькими симметриями и стреловидные конфигурации.
Мы также рассматриваем множества с большим количеством общих точек, которые нельзя объединить.
Эти конструкции подчёркивают потенциал последовательных растяжений и ограничения на объединение множеств,
демонсnрируя новые особенности структуры и свойств плоских целоудалённых множеств.
